---
title: "Tidyverse Create Assignment Vignette"
author: "Chinedu Onyeka"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tidyverse Create Assignment Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Vignette Info  
This vignette demonstrates how to use some functions from the dplyr and ggplot2 packages in the tidyverse suit of packages <br/>  

## Data  
The data used to demonstrate the usage of the dplyr and ggplot2 package was obtained from kaggle. See the link below: <br/>
[Heart Failure Dataset](https://www.kaggle.com/fedesoriano/heart-failure-prediction)

## Load Libraries  

```{r message=FALSE}
library(tidyverse)
```

## Read the data  
```{r read-data, message=FALSE}
url <- "https://raw.githubusercontent.com/chinedu2301/DATA607-Data-Acquisition-and-Management/main/heart.csv"
heart_failure <- read_csv(url)
```
```{r head}
# Look into the data to know what columns are present
head(heart_failure)
```
## dplyr package

### *glimpse*  
Q: What are the variables and data types of the variables in my dataset?  
A: Use dplyr's glimpse function.
```{r}
glimpse(heart_failure)
```

### *select*  
Q: I do not need all the columns, how do I select the columns that I only need for my analysis?  
A: Use dplyr's select function.  
```{r}
# select all columns except Oldpeak and ST_Slope
heart_failure_sub <- heart_failure %>% select(-c(Oldpeak, ST_Slope))
# select columns Age, Sex, RestingBP, and HeartDisease
heart_failure_sub1 <- heart_failure %>% select(Age, Sex, RestingBP, HeartDisease)
# select from columns Age to ExerciseAngina
heart_failure_sub2 <- heart_failure %>% select(Age:ExerciseAngina)
```

View the head of heart_failure_sub
```{r}
head(heart_failure_sub)
```
View the head of heart_failure_sub1
```{r}
head(heart_failure_sub1)
```
View the head of heart_failure_sub2
```{r}
head(heart_failure_sub2)
```
### *filter*  
Q: I want only observations where Sex is M  
A: Use dplyr's filter function.
```{r}
heart_failure_male <- heart_failure %>% filter(Sex == "M")
head(heart_failure_male)
```
Q: I want only observations where RestingECG is Normal  
A: Use dplyr's filter function.  
```{r}
heart_failure_disease <- heart_failure %>% filter(RestingECG == "Normal")
head(heart_failure_disease)
```
## *rename*  
Q: How do I rename a column?  
A: Use dplyr's rename function
```{r}
heart_failure <- heart_failure %>% rename("Gender" = Sex)
head(heart_failure)
```

## ggplot2 package  

### *Scatterplot*  
Q: I want to plot a scatter plot of RestingBP vs. Cholesterol  
A: Use the ggplot function and the geom_point layer  
```{r}
ggplot(data = heart_failure, aes(x = Cholesterol, y = RestingBP)) + geom_point()
```

Add title, transparency, and theme
```{r}
ggplot(data = heart_failure, aes(x = Cholesterol, y = RestingBP)) + geom_point(alpha = 0.5) + 
  labs(title = "RestingBP vs. Cholesterol") + theme_bw()
```

### *Barplot*  
Q: I want to plot a barchart of the count of Male and Female in the dataset  
A: Use the ggplot function and the geom_bar layer
```{r}
ggplot(data = heart_failure, aes(x = Gender)) + geom_bar(fill = "brown") + 
  labs(title = "Bar chart for count of Sex") + theme_bw()
```

Q: How do I flip co-ordinates?  
A: Add the coord_flip layer
```{r}
ggplot(data = heart_failure, aes(x = Gender)) + geom_bar(fill = "brown") + 
  labs(title = "Bar chart for count of Sex") + theme_bw() + coord_flip()
```


















