---
title: "CUNY SPS DATA607 Tidyverse Extend Vignette"
author: "Chinedu Onyeka"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CUNY SPS DATA607 Tidyverse Extend Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 3,
  collapse = TRUE,
  comment = "#>"
)
```
The original Tidyverse assignment was done by [Coffy](https://rpubs.com/blesned/vignette). Code chunks that contains extensions/additions are wrapped with comments in five hash tags.  

Source: [Kagle data - Data Analysis Jobs, Based on NYC Jobs - October 2021](https://www.kaggle.com/intelai/data-analysis-jobs?select=data_analysis_jobs.csv)

This dataset contains current job postings available on the City of New York’s official jobs site ( http://www.nyc.gov/html/careers/html/search/search.shtml ). Internal postings available to city employees and external postings available to the general public are included.


## Load Packages
```{r, message=FALSE}
##### Added 'message=FALSE' in the r chunk to suppress output from loading the libraries
library("tidyverse")
library("reactable")
#####
```



## Data Set

The `readr package` in tidyverse library contains the function `read_csv` that will import data from a csv file. The csv file was downloaded from Kaggle.com dataset on Data Analysis Jobs in NYC for October 2021. The imported data was read in to RStudio as a dataframe, `da`.  
```{r, message=FALSE}
##### Added 'message=FALSE' in the r chunk to suppress output from loading the libraries
da <- read_csv("https://github.com/candrewxs/Vignettes/blob/master/dadata/data_analysis_jobs.csv?raw=true")
##### 
```

```{r}
##### Check the head of the data
head(da)
#####
```




Retrieve the names attribute of the da data set with the `names()` function.
```{r}
names(da)
```
```{r}
##### Check the variables and structure of each variable using dplyr's glimpse function
glimpse(da)
#####
```


## Group by Variables

### Dataset Question: How many job postings are Internal and External? What are the Levels and how many are available?

Create a new data frame using existing data frame (da) by extracting columns: Level and Posting Type. The `count()` combines `group_by` and `count rows in each group` into a single function. Renamed the column name "n" to "Count" and created an interactive data table. 
```{r}
plot1 <- da %>%
   count(Level, `Posting Type`)  # dplyr package: group and count 

colnames(plot1)[3] <- "Count" 
# use function colnames() to rename column 
# and access individual column names with colnames(df)[index]

reactable(plot1) # interactive data table
```

### Visualization 
```{r}
##### Add title and theme to the plot
# discrete visualization with ggplot2
p1 <- ggplot(data = plot1) 
  p1 + geom_col(aes(Level, Count, fill = `Posting Type`)) + labs(title = "Bar Chart of Levels") + theme_bw()
```


```{r}
##### Plot Barchart for only External Posting Type
# discrete visualization with ggplot2
p1 <-  plot1 |> filter(`Posting Type` == "External") |> ggplot(aes(Level, Count))
  p1 + geom_col(fill = "brown") + labs(title = "Bar Chart of Levels -  External") + theme_bw()
#####
```
```{r}
##### Plot Barchart for only Internal Posting Type
# discrete visualization with ggplot2
p1 <-  plot1 |> filter(`Posting Type` == "Internal") |> ggplot(aes(Level, Count))
  p1 + geom_col(fill = "blue") + labs(title = "Bar Chart of Levels - Internal") + theme_bw()
#####
```



### Dataset Question: What are the Business Titles categories which NYC hires Data Analysis to perform? Which Business Title has the highest demand for Data Analysis?

Create a new data frame using existing data frame (da) by extracting column: Business Title. Tidy data frame column to upper case and build a contingency table of the counts at each combination of factor levels with the `table()` function. Renamed the column name "pl2" to "Business Title" and created an interactive data table showing the frequency. 
```{r}
plot2 <- as.data.frame(da[,5])

plot2$`Business Title` <- toupper(plot2$`Business Title`)

pl2 <- as.data.frame(table(plot2))

colnames(pl2)[1] <- "Business Title"

reactable(pl2)
```


```{r}
summary(pl2) # calculates some statistics on the data
```

Filter Business Titles with a frequency greater or equal to 10.  
```{r}
pl_2 <- filter(pl2, Freq >= 10) 

pl_2
```

### Visualization 
```{r}
##### Add theme, title, and color to the bar chart
p2 <- ggplot(pl_2, aes(`Business Title`, Freq)) 
  p2 + geom_col(fill = "brown") + coord_flip() + 
    labs(title = "Frequency of each Job title") + xlab(NULL) + ylab("Frequency") + theme_bw()
#####
```

## Conclusion
Kagle data - Data Analysis Jobs, Based on NYC Jobs was loaded and analyzed using R base functions and packages from Tidyverse and Reactable. These are the Tidyverse packages that were utilized to load `(readr)` , perform data manipulation `(dplyr)` , create graphic representation of input data `(ggplot2)`.



Links
[Rpubs](https://rpubs.com/chinedu2301/833708)

