---
title: "TidyVerse Create Vignette"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

### Title: CUNY SPS MDS DATA607_Tyderverse Create"
### Author: Charles Ugiagbe
### Date: "10/23/2021"


```{r global_options, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
div.aquamarine { background-color:#d9eff9; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">


## Introduction

Tidyverse is just a collection of R packages underlying same design philosophy, grammar, and data structure. There are currently 8 packages in the `tidyverse` package bundle including:

  * `dplyr`: a set of tools for efficiently manipulating datasets;
  * `forcats`: a package for manipulating categorical variables / factors;
  * `ggplots2`: a classic package for data visualization;
  * `purrr`: another set of tools for manipulating datasets, specially vecters, a complement to `dplyr`;
  * `readr`: a set of faster and more user friendly functions to read data than R default functions;
  * `stringr`: a package for common string operations;
  * `tibble`：a package for reimagining data.frames in a modern way;
  * `tidyr`: a package for reshaping data, a complement to `dplyr`.

</div> \hfill\break  
  
**In this assignment, I will use some handy functions in tidyverse package to perform some Analysis**


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

##       Dataset

The dataset in this project is called “student performance” from https://www.kaggle.com/datasets; The dataset contains a sample of 1000 observations of 8 variables.


I use `read.csv` function to import the csv file to R.

```{r}
url <- "https://raw.githubusercontent.com/omocharly/DATA607_PROJECTS/main/StudentsPerformance.csv"
data <- read.csv(url, header = TRUE)
```


```{r}
head(data)
```

Glimpse help us to catch sight of the data to see the data structure.
```{r}
glimpse(data)
```

##   dplyr::rename()

rename() changes the names of individual variables using in a column with a new one


```{r}
data1 <- data %>% rename(race = race.ethnicity, parental_Educatn_level= parental.level.of.education, test.prep = test.preparation.course)
head(data1)
```

##   dplyr::select()

Select(): is use for selecting a range of consecutive variables or taking the complement of a set of variables

```{r}
data2 <- data1 %>%
  select(gender, math.score, reading.score, writing.score)
head(data2)
```

##   dplyr::filter()

I use the filter() function to filter maths, writing and reading scores that are greater than 97

```{r}
data3 <- data2 %>%
  filter(math.score == 100, writing.score > 95, reading.score > 95)
data3
```

##    dplyr::arrange()

arrange(): orders the rows of a data frame by the values of selected columns.

```{r}
data4 <- data2 %>% arrange(desc(math.score))
head(data4)
```

##    dplyr::mutate() 

mutate() adds new variables that are function of the existing ones to the table and also preserves existing ones.

```{r}
data5 <- data4 %>%
  mutate(avg.score = (math.score + writing.score + reading.score) / 3)
head(data5)
```

##       dplyr::case_when() 

case_when: function allows you to vectorise multiple if_else() statements. It is an R equivalent of the SQL CASE WHEN statement. If no cases match

```{r}
data6 <- data5 %>% 
  mutate(pass_fail_grade = case_when(avg.score >= 85 ~ 'Pass'
                                     ,TRUE ~ 'Fail' )
         )
head(data6)
```

##    dplyr::summarize() 

```{r}
data %>% group_by(gender) %>%
  summarize( math_score = sum (math.score)/ n())
```

##    Tidyverse::ggplot2()

ggplot2 is a system for 'declaratively' creating graphics, based on "The Grammar of Graphics".

```{r}
ggplot(data = data6, aes(x = gender, y = avg.score, col = gender), col = red) + geom_boxplot() + labs(title="Distribution of Students Average score") + theme(plot.title = element_text(hjust=0.5))
```

**Other usage of Tidyverse can be found in the textbook "R for Data Science" and other online resource.**
