---
title: "Examples of tidyverse"
author: "Tom Buonora and ..."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tidyverse Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
   warning = FALSE, 
  message = FALSE,
  fig.width = 8,
  fig.height = 8
)
```

```{r}

library(tidyverse)         # ggplot2, dplyr, tidyr, readr, tibble, sringr and more

```


# Tidyverse Vignette




##   readr : read_csv()

<br>
read_csv is part of readr whereas read.csv is base R. Im not sure that read_csv is tidier than read.csv
<br>

```{r readr : read_csv}
CURR_PATH<-str_trim(getwd())


# to do : use the kaggle api
# https://www.kaggle.com/mysarahmadbhat/nyc-traffic-accidents

# nyc_traffic_accidents<-read.csv("NYC Accidents 2020.csv")

nyc_traffic_accidents<-read_csv("https://raw.githubusercontent.com/TheReallyBigApple/CunyAssignments/main/DATA607/NYC%20Accidents%202020.csv",show_col_types = FALSE)


```

##   readr : spec()

the spec function displays the schema for the data frame or tibble

```{r readr : spec}

# the readr spec method shows you the schema
spec(nyc_traffic_accidents)


# this is base R, it removes the spaces from the column names
colnames(nyc_traffic_accidents) <- make.names(colnames(nyc_traffic_accidents))

```

##   dplyr : group_by()    dplyr_summarise()     magrittr : %>%

magrittr provides several operators including the commonly used %>% which pipes results from one function to the next

```{r several}


nyc_traffic_accidents %>%
  group_by(BOROUGH) %>%
  summarise(observations = n(), killed = sum(NUMBER.OF.PERSONS.KILLED, na.rm = TRUE))


```

##   dplyr : select()

select can select certain columns
select can also omit certain columns. 
Below I omit a sequence of columns starting with LATTITUDE and ending with LOCATION

```{r dplyr select}

nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    select(!(LATITUDE:LOCATION))

```

##   dplyr : muatate()   replace_na()

muate() can create or modify a column
replace_na() can replace all instances of NA with a specified value

```{r replace_na}

nyc_traffic_accidents<-nyc_traffic_accidents %>% dplyr::mutate(BOROUGH = replace_na(BOROUGH, "NYC"))

```


##   ggplot :  geom_histogram() et all.

while the other tidyverse packages infer data parsing, ggplot exists to *display* data
having said that, its robust functionality includes data shaping. The histogram is essentially a group_by()
as is scale_x_dates

```{r ggplot histogram}


ggplot(data = nyc_traffic_accidents, aes(x = CRASH.DATE)) +
geom_histogram() +
theme(axis.text.x=element_text(angle=60, hjust=1)) +
ggtitle("Accidents Per Day") +
scale_x_date(date_breaks = "weeks", date_labels = "%Y-%m-%d") +
xlab("Date") 



```


