---
title: "tidyverse: using case_when and str_replace for variable derivation"
author: "William Aiken"
date: "10/24/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(reshape2)
library(kableExtra)
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

## Overview
### This is an example of using the dplyr case_when() function instead of nested ifelse() funcitons.

The data comes from the *fivethirtyeight* article "The Ultimate Halloween Candy Power Ranking"

[link to The Ulitmate Halloween Candy Power Ranking](https://fivethirtyeight.com/videos/the-ultimate-halloween-candy-power-ranking/)

### This is a dataset of 85 common Halloween treats

We read in the data from GitHub using the tidyverse readr package

```{r}
#Read in the data
file_path <- "https://raw.githubusercontent.com/catfoodlover/Data607/main/Data607_HW1_Data_WilliamAiken.csv"

candy_df <- read_csv(file_path, show_col_types = FALSE)

#Select columns of interest and fix names when necessary
candy_df <- candy_df %>% select(names = competitorname, choco = chocolate, fruity, caramel, nuts = peanutyalmondy, nougat, crispy = crispedricewafer, win_percent = winpercent)

#Character needs to be fixed
candy_df$names <- gsub( "Õ", "'", candy_df$names)

# #Convert dataset from wide to long to make it easier to work with
# temp <- melt(data = candy_df, id.vars = "names", measure.vars = c("choco", "fruity", "caramel", "nuts", "nougat", "crispy"), variable.name = "Property", value.name = "Status")

# #Join win percentage back in
# temp2 <- left_join(temp, candy_df %>% select(names, win_percent), by = "names")
```

There are instances where we want to use logic to either update current fields or derive new fields based on a logic evaluation.  In base R, we could use the ifelse() function but if there are multiple conditions we end up with nested ifelse() statements which can be challenging to validate and debug.

With the tidyverse dplyr library we get the case_when function that allows for multiple conditions to be validates at the same time.

In the current data set we have multiple candy properties and we would like to create new categories based on the current properties

We don't have to create exhaustive categories, we can create a fall through case that catches all other values

```{r}
temp <- candy_df %>% mutate(
  choco_composite = case_when(
    choco == 1 &
      nuts == 1 &
      nougat == 1 &
      caramel == 1 ~ "choco_caramel_nougat_nut",
    choco == 1 &
      nuts == 1 &
      nougat == 1 &
      caramel == 0 ~ "choco_nougat_nut",
    choco == 1 &
      nuts == 1 &
      nougat == 0 &
      caramel == 1 ~ "choco_caramel_nut",
    choco == 1 &
      nuts == 1 &
      nougat == 0 &
      caramel == 0 ~ "choco_nut",
    choco == 1 &
      nuts == 0 &
      nougat == 0 &
      caramel == 0 ~ "choco_only",
    choco == 1  ~ "choco_other",
    TRUE ~ "other"
  )
) %>% group_by(choco_composite) %>% summarise(average_win = round(mean(win_percent), 2), count_n = n())

temp %>% select(choco_composite, average_win, count_n) %>% kbl() %>% kable_styling()
```

The case_when() function also allows you to set values to NA if you don't include a fall through case

Here we are going to turn all non-chocolate candy to NA and find the average win percentage

```{r}
temp <- candy_df %>% mutate(
  choco_composite = case_when(
    choco == 1 &
      nuts == 1 &
      nougat == 1 &
      caramel == 1 ~ "choco_caramel_nougat_nut",
    choco == 1 &
      nuts == 1 &
      nougat == 1 &
      caramel == 0 ~ "choco_nougat_nut",
    choco == 1 &
      nuts == 1 &
      nougat == 0 &
      caramel == 1 ~ "choco_caramel_nut",
    choco == 1 &
      nuts == 1 &
      nougat == 0 &
      caramel == 0 ~ "choco_nut",
    choco == 1 &
      nuts == 0 &
      nougat == 0 &
      caramel == 0 ~ "choco_only",
    choco == 1  ~ "choco_other"
  )
)

temp %>% count(choco_composite) %>% kbl() %>% kable_styling()
```

We can also use the tidyverse stringr package to update our 'choco_composite' column.

Say we want to make 'nut' more descriptive.  We can use str_replace() to change all the strings containing 'nut' to 'almond_peanut'

```{r}
temp2 <- temp %>% mutate(choco_composite_update = str_replace(choco_composite, "nut", "almond_peanut"))

temp2 %>% count(choco_composite, choco_composite_update) %>% kbl() %>% kable_styling()
```
