---
title: 'tidyverse: using purrr to manipulate dataframe'
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Tidyverse: Using `purrr` to manipulate data in a functional way

## Load Tidyverse
```{r results=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Dataset

Data was obtained from FiveThirtyEight's article (How Americans View Biden's Response To The Coronavirus Crisis)[https://projects.fivethirtyeight.com/coronavirus-polls/]

```{r, warning=FALSE, message=FALSE}
approvalPollsDf <- read_csv("https://raw.githubusercontent.com/nolivercuny/data607/master/covid_approval_polls.csv")

glimpse(approvalPollsDf)
```

## Using `map2`

Q: What is the difference between approval and disapproval

A: Use the `map2` function to apply a difference function across the approval and disapproval columns


```{r}
approvalPollsDf$approve_dissapprove_diff <- map2_dbl(approvalPollsDf$approve, approvalPollsDf$disapprove, ~{.x - .y})
approvalPollsDf %>% head() %>% glimpse()
```

## Using `some`

Q: Can we use `purrr` to check if our new column contains any `NA` values?

A: Yes, you can use the `some` function which returns `TRUE` if at least one value matches your condition

```{r}
some(approvalPollsDf$approve_dissapprove_diff, is.na)
```
 
## Using `modify_if`

Q: How can we coerce the `NA` values into a number

A: You can use the `modify_if` function which will modify a value only if the predicate returns true.

```{r}
approvalPollsDf$approve_dissapprove_diff <- modify_if(approvalPollsDf$approve_dissapprove_diff, is.na, ~0.0)
```

Using it again to add a new column marking if the majority approve or disapprove

```{r}
approvalPollsDf$majority_approve <-
map_if(approvalPollsDf$approve_dissapprove_diff, ~.x > 0, ~"TRUE", .else = ~"FALSE") %>% unlist()
```

## Plot the data

Box plot comparing the approval and disapproval margins. I use the `map_dbl` to convert the negative values to positive for side-by-side comparison

```{r}
approvalPollsDf %>% ggplot(aes(x=majority_approve,y=map_dbl(approve_dissapprove_diff,abs),fill=majority_approve)) +
  geom_boxplot()
```

