---
title: "tidyverse_create vignette"
author: "Javier Pajuelo Bazan"
date: "11/4/2021"
output:
 prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## OVERVIEW
TidyVerse CREATE assignment

Your task here is to Create an Example.  Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle,
create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse
package with your selected dataset. (25 points)

We are using the kaggle URL:
https://www.kaggle.com/uciml/red-wine-quality-cortez-et-al-2009?select=winequality-red.csv

```{r ,message=FALSE, echo=TRUE}
WINE_QUALITY_RAW_URL <- 'https://raw.githubusercontent.com/quaere1verum/sps_public/master/data607-001/assignments/tidyverse_assignments/winequality-red.csv'

library(curl)
# using tidy verse packages
# -- Attaching packages -------------------------------------------------------------------------------------------------- tidyverse 1.3.1 --
#   v ggplot2 3.3.5     v purrr   0.3.4
# v tibble  3.1.3     v stringr 1.4.0
# v tidyr   1.1.3     v forcats 0.5.1
# v readr   2.0.1     
library(tidyverse)            
```


## Using readr package from tidyverse
Readr package contains the read_csv function required to load our data.

```{r ,message=FALSE, echo=TRUE}
load_csv_from_url <- function(url_path)
{
  tmp <- tempfile()
  curl_download(url_path, tmp)
  read_csv(tmp)               # read_csv() requires dependencies from tidyverse, readr package 
}

wine_quality_data <- load_csv_from_url(WINE_QUALITY_RAW_URL)
```

## Using purr package from tidyverse
Purr package has functions like maps data to functions and returns vectors.

```{r ,message=FALSE, echo=TRUE}
# using purr package from tidyverse
wine_quality_data %>%
  split(.$alcohol) %>% # from base R
  map(~ lm(`volatile acidity` ~ `citric acid`, data = .)) %>%
  map(summary) %>%
  map_dbl("r.squared")  # map_dbl is a purr function returns double vectors

```

## Using gplot package from tidyverse
Adding a few cool plots that show some kind of relationship between acidity and alcohol quantity and quality.

```{r ,message=FALSE, echo=TRUE}
ggplot(wine_quality_data, aes(`fixed acidity`, `volatile acidity`, colour = alcohol)) + 
  geom_point()
```

```{r ,message=FALSE, echo=TRUE}
ggplot(wine_quality_data, aes(`fixed acidity`, `volatile acidity`, colour = quality)) + 
  geom_point()


# Later, you'll be asked to extend an existing vignette.  Using one of your classmate’s examples (as created above), you'll then 
# extend his or her example with additional annotated code. (15 points)
# I think this 'Later' reference means to the tidy verse extend assignment

##### Javier's vignette ends #####
```


