---
title: "Untitled"
author: "Santiago Torres"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(purrr)
```

# Purr functions for looping through data

## Comparing followers of world religions

Source: [FiveThirtyEight](https://github.com/fivethirtyeight/data/tree/master/nfl-elo)

I was looking for a dataset where I could use the functions within `purrr`

```{r, warning=FALSE}
nfl_elo = read_csv("https://raw.githubusercontent.com/sserrot/DATA607/main/Tidyverse%20Project/nfl-elo/nfl_elo_latest.csv") 

glimpse(nfl_elo)
```

## 1A: basic reduce function

Q: What is the sum of all the scores in the data set??
A: Use `reduce` to combine a vector

```{r}
non_na_scores_1 <- nfl_elo$score1[!is.na(nfl_elo$score1)]
non_na_scores_2 <- nfl_elo$score2[!is.na(nfl_elo$score2)]
non_na_scores_1

non_na_scores_1 %>% reduce(`+`)
non_na_scores_2  %>% reduce(`+`)
```
 
## 1B: basic map function

Q: How many total points were scored in a game??  
A: Use `map2` to apply a function across a two vectors

```{r}
map2(non_na_scores_1, non_na_scores_2, sum)
```
