---
title: "Examples of tidyverse"
author: "Tom Buonora and Shane Hylton -- EXTEND"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tidyverse Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
   warning = FALSE, 
  message = FALSE,
  fig.width = 8,
  fig.height = 8
)
```


```{r}

library(tidyverse)         # ggplot2, dplyr, tidyr, readr, tibble, sringr and more

```


# Tidyverse Vignette



## readr : read_csv()

<br>
read_csv is part of readr whereas read.csv is base R. I'm not sure that read_csv is tidier than read.csv
<br>

```{r readr : read_csv}
CURR_PATH<-str_trim(getwd())


# to do : use the kaggle api
# https://www.kaggle.com/mysarahmadbhat/nyc-traffic-accidents

# nyc_traffic_accidents<-read.csv("NYC Accidents 2020.csv")

nyc_traffic_accidents <- read_csv("https://raw.githubusercontent.com/TheReallyBigApple/CunyAssignments/main/DATA607/NYC%20Accidents%202020.csv",show_col_types = FALSE)


```

##   readr : spec()

the spec function displays the schema for the data frame or tibble

```{r readr : spec}

# the readr spec method shows you the schema
spec(nyc_traffic_accidents)


# this is base R, it removes the spaces from the column names
colnames(nyc_traffic_accidents) <- make.names(colnames(nyc_traffic_accidents))

```

##   dplyr : group_by()    dplyr_summarise()     magrittr : %>%

magrittr provides several operators including the commonly used %>% which pipes results from one function to the next

```{r several}

nyc_traffic_accidents %>%
  group_by(BOROUGH) %>%
  summarise(observations = n(), killed = sum(NUMBER.OF.PERSONS.KILLED, na.rm = TRUE))

```
### An Extension of group_by( ) using arrange ( )

Here, we can group by borough and rearrange the dataframe to make viewing the data by borough simpler. 

```{r}

nyc_traffic_accidents <- nyc_traffic_accidents %>%
  group_by(BOROUGH) %>%
  arrange(BOROUGH)

```


##   dplyr : select()

select can select certain columns
select can also omit certain columns. 
Below I omit a sequence of columns starting with LATTITUDE and ending with LOCATION

```{r dplyr select}

nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    select(!(LATITUDE:LOCATION))

```

### dplyr::relocate( ) -- Extension 

We can further this tidying of the dataframe by rearranging the data how we see fit. 

```{r}

nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    relocate(BOROUGH, .before = CRASH.DATE)

nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    relocate(VEHICLE.TYPE.CODE.1, .after = BOROUGH)
nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    relocate(VEHICLE.TYPE.CODE.2, .after = VEHICLE.TYPE.CODE.1)

nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    relocate(CONTRIBUTING.FACTOR.VEHICLE.1, .after = VEHICLE.TYPE.CODE.2)
nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    relocate(CONTRIBUTING.FACTOR.VEHICLE.2, .after = CONTRIBUTING.FACTOR.VEHICLE.1)

nyc_traffic_accidents[1:5,1:5]

```

Here, the data was rearranged to show the types of vehicles involved and the reasons for the crash at the beginning of the dataframe instead of towards the end. 


### An Extension of group_by ( ) and summarize ( )

Another use for summarize would be to show which zip codes had the most injuries. This can be further extended to a plot to visualize the injuries by zip code. Because there are nearly 200 different zip codes, it may be challenging to plot them all together.

```{r}

zip_injuries <- nyc_traffic_accidents %>%
  group_by(ZIP.CODE) %>%
  summarize(injuries = sum(NUMBER.OF.PERSONS.INJURED, na.rm = TRUE))

zip_injuries

```

##   dplyr : muatate()   replace_na()

muate() can create or modify a column
replace_na() can replace all instances of NA with a specified value

```{r replace_na}

nyc_traffic_accidents<-nyc_traffic_accidents %>% dplyr::mutate(BOROUGH = replace_na(BOROUGH, "NYC"))

```


##   ggplot :  geom_histogram() et all.

while the other tidyverse packages infer data parsing, ggplot exists to *display* data
having said that, its robust functionality includes data shaping. The histogram is essentially a group_by()
as is scale_x_dates

```{r ggplot histogram}


ggplot(data = nyc_traffic_accidents, aes(x = CRASH.DATE)) +
geom_histogram() +
theme(axis.text.x=element_text(angle=60, hjust=1)) +
ggtitle("Accidents Per Day") +
scale_x_date(date_breaks = "weeks", date_labels = "%Y-%m-%d") +
xlab("Date") 



```

### An Extension of ggplot ( )

First, I will show a pitfall of too much or improperly grouped data:

```{r}

ggplot(zip_injuries, aes(x=ZIP.CODE, y = injuries))+
  geom_col()+
  #geom_density(alpha=.2, fill = 'red')+
  ggtitle('Injuries by ZIP Code')+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab('ZIP Code')+
  ylab('Injuries')

```

There are a few issues here. First, with the x representing ZIP.CODE, ggplot will sort the x-axis data based on the x-values, which are five-digit numbers that should not have any order in this plot. Second, there are nearly 200 entries to work with, which leads to a very busy plot. 

To remedy this, we need a smaller subset of the zip code injuries dataframe. For simplicity, I will plot the top ten as a barplot. geom_col( ) and geom_bar( ) are very similar with some slightly different rules. The main difference is that geom_bar( ) uses stat_count( ) for y values while geom_col( ) uses stat_identity( ) for yy values. Another good solution is to sort the bars in descending order. This can be accomplished by arranging the data in descending order of injuries and then plotting using the index as the x-value. 

```{r}

zip_injuries <- zip_injuries %>% arrange(desc(injuries))
zip10 <- head(zip_injuries, n = 10)
zip10 <- na.omit(zip10)

zip10$idx <- as.numeric(row.names(zip10))

ggplot(zip10, aes(x=idx, y = injuries, color = ZIP.CODE))+
  geom_col(aes(fill = ZIP.CODE))+
  #geom_density(alpha=.2, fill = 'red')+
  ggtitle('Injuries by ZIP Code')+
  theme(plot.title = element_text(hjust = 0.5, size = 20))+
  theme(text = element_text(size = 15))+
  xlab('ZIP Code')+
  geom_text(
    label=zip10$ZIP.CODE, 
    nudge_x = 0, nudge_y = 10, 
    check_overlap = T,
    size = 5
  )+
  ylab('Injuries')

```



