---
title: "Extend from TidyVerse scatter-plot options of CLAUDIO, Mauricio"
extended by: 'Trang Do'
date: 'Due Date: 2021-11-07'
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
vignette: |
  %\VignetteIndexEntry{Vignette Title} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

Using [Kaggle's Countries of the World data set](https://www.kaggle.com/fernandol/countries-of-the-world), we'll explore some ggplot scatter-plot options.  

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(stringr)
#cotw = read.csv("C:\\Users\\LENOVO\\OneDrive\\Learning\\Courses In Progress\\DATA 607\\TidyVerse CREATE\\countries of the world.csv")
cotw=read.csv("https://raw.githubusercontent.com/MauricioClaudio/CUNY-MPS-DATA-607/main/countries%20of%20the%20world.csv")
cotw$Infant.mortality..per.1000.births.=str_replace(cotw$Infant.mortality..per.1000.births., ",",".")
cotw$Infant.mortality..per.1000.births.=as.numeric(cotw$Infant.mortality..per.1000.births.)
```


## Basic scatter plot

```{r, warning=FALSE, message=FALSE}
ggplot(cotw, ## Specify data source
       aes(x=GDP....per.capita., ## Specify x-axis variable
                y=Infant.mortality..per.1000.births.)) + ## Specify y-axis variable
  geom_point() + ## Specifiy 'scatter point' plot
  xlab("GDP / capita") + ## Specify x-axis label
  ylab("Infant Mortality / 1000 births") ## Specify y-axis label
```

## Add logarithmic scale to the x-axis

Add a ``scale_x_log10()`` call.

```{r, warning=FALSE}
ggplot(cotw, ## Specify data source
       aes(x=GDP....per.capita., ## Specify x-axis variable
                y=Infant.mortality..per.1000.births.)) + ## Specify y-axis variable
  geom_point() + ## Specifiy 'scatter point' plot
  xlab("GDP / capita") + ## Specify x-axis label
  ylab("Infant Mortality / 1000 births") + ## Specify y-axis label
  scale_x_log10() ## Add logarithmic scale on the x-axis
```

## Add a third variable to the plot with dot color  

Assign a nominal variable to *color* in ``aes()``.  

```{r, warning=FALSE}
ggplot(cotw, ## Specify data source
       aes(x = GDP....per.capita., ## Specify x-axis variable
                y = Infant.mortality..per.1000.births., ## Specify y-axis variable
                color = Region  ## Specify color variable
           )) + 
  geom_point() + ## Specifiy 'scatter point' plot
  xlab("GDP / capita") + ## Specify x-axis label
  ylab("Infant Mortality / 1000 births") + ## Specify y-axis label
  scale_x_log10() ## Add logarithmic scale on the x-axis
```

 

## Add a fourth variable to the plot with dot size  

Assign a numerical variable to *size* in ``aes()``.

```{r, warning=FALSE}
ggplot(cotw, ## Specify data source
       aes(x = GDP....per.capita., ## Specify x-axis variable
                y = Infant.mortality..per.1000.births., ## Specify y-axis variable
                color = Region,  ## Specify color variable
                size = Population  ## Specify size variable
           )) + 
  geom_point() + ## Specifiy 'scatter point' plot
  xlab("GDP / capita") + ## Specify x-axis label
  ylab("Infant Mortality / 1000 births") + ## Specify y-axis label
  scale_x_log10() ## Add logarithmic scale on the x-axis
```

## Decrease the opacity of the dots

Specify *alpha* parameter in ``geom_point()``.

```{r, warning=FALSE}
ggplot(cotw, ## Specify data source
       aes(x = GDP....per.capita., ## Specify x-axis variable
                y = Infant.mortality..per.1000.births., ## Specify y-axis variable
                color = Region,  ## Specify color variable
                size = Population  ## Specify size variable
           )) + 
  geom_point(alpha=1/3) + ## Specifiy 'scatter point' plot & dot opacity
  xlab("GDP / capita") + ## Specify x-axis label
  ylab("Infant Mortality / 1000 births") + ## Specify y-axis label
  scale_x_log10() ## Add logarithmic scale on the x-axis
```


## Partition or facet into several plots  

Add ``facet_wrap(~ x)``, where *x* is the partition variable.

```{r, warning=FALSE}
ggplot(cotw, ## Specify data source
       aes(x = GDP....per.capita., ## Specify x-axis variable
                y = Infant.mortality..per.1000.births., ## Specify y-axis variable
                color = Region,  ## Specify color variable
                size = Population  ## Specify size variable
           )) + 
  geom_point() + ## Specifiy 'scatter point' plot
  xlab("GDP / capita") + ## Specify x-axis label
  ylab("Infant Mortality / 1000 births") + ## Specify y-axis label
  scale_x_log10() + ## Add logarithmic scale on the x-axis
  facet_wrap(~ Region) +  ## Specify the partition/facet variable
  theme(strip.text.x = element_blank(),
        legend.position = "bottom",
        legend.box="vertical", 
        legend.margin=margin(),
        legend.text = element_text(size=5.5),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Coordinate data with World Map (Extend)

geom_map to plot & iso3166 to get country name & country code

```{r Extend}
library(dplyr)
library(maps)
library(stringr)
library(rworldmap)
library(ggplot2)
library(sqldf)

dat <- iso3166
head(dat)

countries <- as.data.frame(dat)
countries  <-rename(countries, Country = ISOname)

cotw<-cotw %>%
  mutate(Country=str_trim(Country),Infantmortality=Infant.mortality..per.1000.births.)

cotw<-merge(countries, cotw, by = "Country")

mapworld <- map_data(map="world")

mapworld <-mapworld %>%
  mutate(Infantmortality = 0,Country=str_trim(region),Abbreviation="")

mapworld <- sqldf(c("UPDATE mapworld
             SET Infantmortality = (SELECT Infantmortality
                          FROM cotw
                          WHERE mapworld.Country = cotw.Country),
                Abbreviation = (SELECT a3
                          FROM cotw
                          WHERE mapworld.Country = cotw.Country)          
             WHERE EXISTS (SELECT 1
                           FROM cotw
                           WHERE mapworld.Country = cotw.Country)"
             , "select * from main.mapworld")
           )
geomtext <- mapworld %>%
  group_by(Abbreviation) %>%
  arrange(Abbreviation) %>%
  filter(row_number()==1)

ggplot()+
   theme(legend.position="none")+
   geom_map(data=mapworld, map=mapworld, aes(map_id=region, x=long, y=lat, fill=Infantmortality))+
  scale_fill_gradient(low = "pink", high = "red", guide = "colourbar")

#  geom_text(data=geomtext,mapping=aes(x=long,y=lat,label=Abbreviation))
```



