---
title: "Kagle Data - Tidyverse"
author: "Coffy Andrews-Guo   Tom Buonora"
date: "`r Sys.Date()`"
output:
  html_document:
    includes:
      in_header: header.html
    toc: true
    toc_float: true
    
    
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 3,
  collapse = TRUE,
  comment = "#>"
)
```


# TidyverseCreate
Source: [Kagle data - Data Analysis Jobs, Based on NYC Jobs - October 2021](https://www.kaggle.com/intelai/data-analysis-jobs?select=data_analysis_jobs.csv)

This dataset contains current job postings available on the City of New York’s official jobs site ( http://www.nyc.gov/html/careers/html/search/search.shtml ). Internal postings available to city employees and external postings available to the general public are included.


## Load Packages
```{r}
library("tidyverse")
library("reactable")
library("tidytext")
```



## Data Set

The `readr package` in tidyverse library contains the function `read_csv` that will import data from a csv file. The csv file was downloaded from Kaggle.com dataset on Data Analysis Jobs in NYC for October 2021. The imported data was read in to RStudio as a dataframe, `da`.  
```{r}
da <- read_csv("https://github.com/candrewxs/Vignettes/blob/master/dadata/data_analysis_jobs.csv?raw=true")
```



Retrieve the names attribute of the da data set with the `names()` function.
```{r}
names(da)
```

## Group by Variables

### Dataset Question: How many job postings are Internal and External? What are the Levels and how many are available?

Create a new data frame using existing data frame (da) by extracting columns: Level and Posting Type. The `count()` combines `group_by` and `count rows in each group` into a single function. Renamed the column name "n" to "Count" and created an interactive data table. 
```{r}
plot1 <- da %>%
   count(Level, `Posting Type`)  # dplyr package: group and count 

colnames(plot1)[3] <- "Count" 
# use function colnames() to rename column 
# and access individual column names with colnames(df)[index]

reactable(plot1) # interactive data table
```

### Visualization 
```{r}
# discrete visualization with ggplot2
p1 <- ggplot(data = plot1) 
  p1 + geom_col(aes(Level, Count, fill = `Posting Type`))
```


### Dataset Question: What are the Business Titles categories which NYC hires Data Analysis to perform? Which Business Title has the highest demand for Data Analysis?

Create a new data frame using existing data frame (da) by extracting column: Business Title. Tidy data frame column to upper case and build a contingency table of the counts at each combination of factor levels with the `table()` function. Renamed the column name "pl2" to "Business Title" and created an interactive data table showing the frequency. 
```{r}
plot2 <- as.data.frame(da[,5])

plot2$`Business Title` <- toupper(plot2$`Business Title`)

pl2 <- as.data.frame(table(plot2))

colnames(pl2)[1] <- "Business Title"

reactable(pl2)
```


```{r}
summary(pl2) # calculates some statistics on the data
```

Filter Business Titles with a frequency greater or equal to 10.  
```{r}
pl_2 <- filter(pl2, Freq >= 10) 

pl_2
```

### Visualization 
```{r}
p2 <- ggplot(pl_2) 
  p2 + geom_col(aes(`Business Title`, Freq)) +
    coord_flip()
```

## Conclusion
Kagle data - Data Analysis Jobs, Based on NYC Jobs was loaded and analyzed using R base functions and packages from Tidyverse and Reactable. These are the Tidyverse packages that were utilized to load `(readr)` , perform data manipulation `(dplyr)` , create graphic representation of input data `(ggplot2)`.



Links
[GitHub](https://rpubs.com/blesned/vignette)



# TidyverseExtend


<br><br>
In the Extend chapter, we will take a look at the *Preferred Skills* field which is a free format english description
of the requirements. We are going to do some lexicon analysis on this field.
<br><br>

Note there are some non ascii characters in it.

```{r}

da[1,]$`Preferred Skills`
```

<br><br>

## stringr::str_replace_all()


<br><br>
To get rid of these we can use the *str_replace_all()* command to replace anything thats not between the hexadecimal codes that encompasses all
ascii characters. Note *str_replace_all()* is the equivalent of the the base R *gsub()* function.
<br><br>

```{r}


da$`Preferred Skills`<-str_replace_all(da$`Preferred Skills`,"[^\x20-\x7E]", "")

```

<br><br>
Note the non-ascci characters are removed.
<br><br>
```{r}
da[1,]$`Preferred Skills`
```
<br><br>


## tidyr::drop_na()

<br><br>
Use *drop_na()* to eliminate any null values.
<br><br>
```{r}

pref_skills_df<-da[c("Job ID","Preferred Skills")]


names(pref_skills_df)<-c("jobid", "skills" )     # set col names

# drop the NA skills
pref_skills_df <- pref_skills_df %>%
    drop_na(skills)               



```
<br><br>


## tidytext::unnest_tokens()


<br><br>
Now we can use *unnest_tokens()* to seperate every word into a seperate row.
The first parameter "word" is the token parameter, so you could specify token=word.
Other tokens include sentence, line, and character.
<br><br>
```{r}

pref_skills_df2 <- pref_skills_df %>%
  unnest_tokens(word, skills )

```

<br><br>

## dplyr :: group_by() and summarize()


<br><br>
These are 2 very useful functions that often go together.
Here we get the count of every word in the job requirements.
<br><br>

```{r}

pref_skills_df3 <- pref_skills_df2 %>%
  group_by(word) %>%
  summarize(n = n())

```
<br><br>

## dplyr :: arrange()


<br><br>
Use *arrange()* to sort by the count, descending, so we can see the most used words.
Note *arange()* is the equivalent to the base R *order()*
<br><br>
```{r}

 pref_skills_df3 %>%
      arrange(desc(n)) %>%
        head()
  
```
<br><br>


## dplyr::filter()

<br><br>
We are only interested in skills so lets create a corpus of skills and use the *filter()* function to isolate them.
<br><br>
```{r}

  

skill_words<-c("management","communication","written","microsoft","excel","analysis","business","sql","interpersonal","software","research","powerpoint","database","quantitative","java","databases","statistical","leadership","communications","javascript","oracle","agile","visio","tableau","law","xml","manager","html5","spanish","economics","github","osha","qa")


pref_skills_df4<-pref_skills_df3 %>%
    filter(word %in% skill_words) %>%
    arrange(desc(n))

```
<br><br>
## dplyr::inner_join()
<br><br>
Now use *inner_join()* to merge the words back to the job ids. Note *inner_join()* is
the equivalent of the base R *merge()* function.
<br><br>
```{r}

pref_skills_df5= pref_skills_df2 %>% inner_join(pref_skills_df4,by="word")

```

<br><br>

## reactable

<br><br>

Lastly, the reactable package is a pretty neat alternative to knitr and kable.
<br><br>
You can set up a filter, for example type "business" under word.
<br><br>
```{r}
reactable(pref_skills_df5, filterable = TRUE, minRows = 10)
```
<br><br>
You can display in groups
<br><br>
```{r}
reactable(pref_skills_df5, groupBy = "word")

```


See [here](https://glin.github.io/reactable/articles/examples.html) for a great tutorial on reactable.