---
title: "Vignette That Demonstrates the functions in TidyVerse"
author: "Nnaemeka Newman Okereafor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Programming Vignette That Demonstrates the functions TidyVerse packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### About the Assignment

In this assignment, I x-rayed the functionality of several TidyVerse packages used in r-programming for data munging, cleaning and visualization. To achieve this, I have written a funtion named *datasciencetools* that takes two parameters namely; data frame and a question. The goal of the *datasciencetools* is to receive a question and returns a *data frame* and a *plot* that contains the count and proportion of data scientist in the United states associated with the question. While building the function, I utilized different functions in **Dplyr, ggplot2 and Tidyr** also known as **TidyVerse packages**

### Data

The dataset is a 2018 multiplechoice interview conducted by Kaggle to gather information on the state of data science and machine learning around the world. The dataset is made up of about 23,859 observations. The dataset is a congregation of responses collected across the globe. The data can be viewed on  [data](https://www.kaggle.com/kaggle/kaggle-survey-2018?select=SurveySchema.csv).

### Function

```{r, warning=FALSE, include=FALSE}
library(tidyverse)
library(stringr)
library(readr)

url<-"https://raw.githubusercontent.com/baruab/msdsrepo/main/Project_3_607/kaggle-survey-2018/multipleChoiceResponses.csv"
survey_tib <- read.csv(file=url(url), sep=",")
```


```{r datasciencetools}

datasciencetools<- function(ds,question){
  names(ds) <- paste(names(ds),ds[1,],sep="_")
  ds <- ds[-c(1),]
  
# filter function is used to tailor the location/country of data scientist to the USA and current role to Data science  
  ds <- ds %>% 
    filter((`Q3_In which country do you currently reside?`=='United States of America')&(`Q6_Select the title most similar to your current role (or most recent title if retired): - Selected Choice`=='Data Scientist'))
# select function is used to select the variable/question from the data frame while the pivot_longer function is used to restructure a wide data set to a long one  
  ds_tool<- ds %>% select(starts_with(question),-contains("OTHER_TEXT")) %>%
    pivot_longer(starts_with(question),names_to="ToolName",values_to="Tool")
  
# count function counts the number of unique values of a variable  
  ds_tool_tib <- ds_tool %>%
    count(ds_tool$Tool) %>% rename("VisualTools" = "ds_tool$Tool","Count"="n")
  
# The arrange function is used to sort the values either in a descending or ascending order  
  ds_tool_tib <- ds_tool_tib%>%filter(!(VisualTools==""))%>%
    arrange(desc(Count))
# mutate function offered the opportunity for a new variable to be created for a dataframe 
  prop_n_count<-ds_tool_tib %>%
    mutate(proportion = round((Count /sum(Count))*100,2))%>% arrange(desc(proportion))
  print(prop_n_count)
# ggplot function and geom_col makes visualization possible for the ananlsis  
  prop_n_count_view <- prop_n_count %>% ggplot(aes(reorder(VisualTools,proportion),proportion)) +
    geom_col()+coord_flip()+geom_col(fill="#A7ADBE")+
    geom_text(aes(label=proportion),color="red") +labs(x="Tools/Technique")+theme_bw()
  
  return(prop_n_count_view)
  
  
}

```


### What types of data do you (Data Scientists) interact with most often at work or school?

```{r}
datasciencetools(survey_tib,"Q31")
```

### Filter Function

In dplyr, filter() function takes in 2 arguments namely; the dataframe you are operating on and a conditional expression that evaluates to TRUE or FALSE. Under the hood, dplyr filter() funtion works by testing each row against your conditional expression and mapping the results to TRUE and FALSE. It then selects all rows that evaluate to TRUE. However, one can use <, >, <=, >=, ==, and != in similar ways to filter any data. The filter() function played two major roles in the  *datasciencetools* function namely; to select the country of interest (USA) and job role from the global survey dataset. Secondly, to filter out missing values from the dataframe.



### What is the favorite media sources that report on data science topics?

```{r}
datasciencetools(survey_tib,"Q38")
```


### Select Function

Select function in r-programming is used to select variables in R using Dplyr package.  Dplyr package in R is provided with select() function which select the columns based on conditions. Select() function in dplyr which is used to select the columns based on conditions like starts with, ends with, contains and matches certain criteria and also selecting column based on position. In the *datasciencetools* function, the select() function alongside with start_with() function and contains() function played a vital role. The select() function used the start_with() function as a parameter to select variables that starts with a particular alphabet.The contains() function with a negative notation parsed in the select() function helped to restrict variables that contains *OTHER_TEXT* from being selected.


### How do you perceive the quality of online learning platforms and in-person bootcamps as compared to the quality of the education provided by traditional brick and mortar institutions? - Online learning platforms and MOOCs:


```{r}
datasciencetools(survey_tib,"Q39")
```


### Count Function

The count() function allows the quickly count of unique values of one or more variables. The *datasciencetools* function utilized the count() to count the unique occurrence of a particular tool or instance of the question in each category.



### Which of the following relational database products have you used at work or school in the last 5 years?

```{r}
datasciencetools(survey_tib,"Q29")
```

### Arrange Function

This is one of the dplyr functions in TidyVerse package. The arrange function is used to sort data based on a specific variable. All we need to do is to make sure that the library is specified and then call the arrange function, passing it the data frame and the variable name.In the *datasciencetools* fuction, the arrange() was used to sort the dataframe by the count variable. The arrange() used the *desc()* function as an argument to sort the dataframe in a decending order.


### Which of the following cloud computing products have you used at work or school in the last 5 years?

```{r}
datasciencetools(survey_tib,"Q27")
```


### GGPLOT Function

GGPLOT function is part of the GGPLOT2 library of the tidyVerse package. It is a visualization function. It has three components which include *Data*, *Geometry* and *Aesthetic mapping*. In *datasciencetools* function, the summarized global data survey data table is the data component. The barplots is the geometry component and the color unicodes pased to the geom_col() function makes up the aesthetics. Other possible geometries are scatter, histogram, smooth densities, and boxplot. However, the plot uses several visual cues to represent the information provided by the dataset. The two most important cues in this plot are the point positions on the x-axis and y-axis. 













