---
title: "TIdyverse Create"
author: "Jiho Kim"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Retrieve Data using readr::read.csv

I chose heart failure dataset from Kaggle and read the csv file using read.csv function.

```{r}
heartdata <- read.csv('https://raw.githubusercontent.com/jihokim97/FALL2021TIDYVERSE/Tidyverse/heart.csv')
view(heartdata)
heartdata
```

## Dplyr

I used dplyr packages to demonstrate different capabilities. 

## Filter

We can use filter functions to filter the desired rows from the dataframe.
```{r}
filteredheart <- heartdata%>%
  filter(Cholesterol > 200)
filteredheart
```

## Select

We can select certain desired columns using select function.

```{r}
selectheart <- heartdata%>%
  select(c('Age','MaxHR','HeartDisease'))
selectheart
```

## group_by and Summarise

We can group dataframe by columns and create new data frame into a single row using summarise.

```{r}
bysex <- heartdata%>%
  group_by(Sex)%>%
  summarise(count = n())

bysex
```
