---
title: "DATA 607 - Tidyverse - Reading and writing your data"
author: "Preston Peck"
date: "10/24/2021"
output: html_document
---

```{r message = FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(googlesheets4)
```

Source: https://github.com/fivethirtyeight/data/tree/master/state-of-the-state

# Read tabular data with readr
##### Most pre-configured parsing methods can be re-written using read_delimeter without anything being lost
```{r}
readDirectory <- "data/read/"
```

## Methods Values separated by anything
### Values separated by anything
##### trim_ws defaults to FALSE unlike other functions, which means blanks won't convert to NA
```{r message = FALSE}
anythingFile <- "words.txt"
anythingPath <- paste(readDirectory, anythingFile, sep = "")

anything <- read_delim(anythingPath, delim = "|", trim_ws = TRUE)
```

### Comma separated values
```{r message = FALSE}
commaFile <- "words_comma.csv"
commaPath <- paste(readDirectory, commaFile, sep = "")

csvComma <- read_csv(commaPath)
csvComma

all.equal(csvComma, read_delim(commaPath, delim = ",", trim_ws = TRUE))
```

### Semi-colon separated values
##### In various European locales, as the comma character serves as the decimal point, the function read_csv2 should be used instead. Decimal points as a result are parsed out, and semi-colons are used in place of commas as delimiters. We can use read_delim with locale to achieve similar results
```{r message = FALSE}
semiColonFile <- "words_semi-colon.csv"
semiColonPath <- paste(readDirectory, semiColonFile, sep = "")

csvSemiColon <- read_csv2(semiColonPath)
csvSemiColon

all.equal(csvComma, read_delim(semiColonPath, delim = ";", trim_ws = TRUE, locale = locale(decimal_mark = ",")))
```

### Tab separated values
```{r message = FALSE}
tabFile <- "words.tsv"
tabPath <- paste(readDirectory, tabFile, sep = "")

tsv <- read_tsv(tabPath)
tsv

all.equal(tsv, read_delim(tabPath, delim = "\t", trim_ws = TRUE))
```

### Whitespace separated values
##### read_table looks for whitespace separation of varying lengths. This makes strings with spaces unable to be parsed
```{r message = FALSE}
table <- read_table(tabPath)
table

all.equal(tsv, table)
```

### Fixed width values
##### Column names can't be extracted from headers, and there are 3 parsing variations:\n
##### fwf_empty() tries to intelligently look for consistent widths, so mileage varies\n
##### fwf_widths() reads continuously according to lengths provided\n
##### fwf_positions() read between indexes, so it can be discontinuous
```{r message = FALSE}
colNames <- c(
  "phrase",
  "category",
  "d_speeches",
  "r_speeches",
  "total",
  "percent_of_d_speeches",
  "percent_of_r_speeches",
  "chi2",
  "pval"
)

fixedWidthFile <- "words_fwf.txt"
fixedWidthPath <- paste(readDirectory, fixedWidthFile, sep = "")

fwfEmpty <- read_fwf(fixedWidthPath)
fwfEmpty

widths <- c(28, 30, 3, 3, 3, 6, 6, 12, 11)

fwfWidth <- read_fwf(fixedWidthPath, fwf_widths(widths, colNames))
fwfWidth

startPositions <- c(1, 29, 59, 62, 65, 68, 74, 80, 92)
endPositions <- c(28, 58, 61, 64, 67, 73, 79, 91, 102)

fwfPositions <- read_fwf(fixedWidthPath, fwf_positions(startPositions, endPositions, colNames))
fwfPositions

all.equal(fwfEmpty, fwfWidth)
all.equal(fwfWidth, fwfPositions)
```

## Compare Methods
##### The data content is identical between methods, but the data frames differ ever so slightly in a hidden delim attribute that will prevent recognizing equality. By ignoring attribute checks, we will just compare content
```{r message = FALSE}
all.equal(anything, csvComma)
attributes(anything)

attributes(anything)$spec$delim
attributes(csvComma)$spec$delim
attributes(csvSemiColon)$spec$delim
attributes(tsv)$spec$delim
attributes(fwfWidth)$spec$delim

all.equal(csvComma, anything, check.attributes = FALSE)
all.equal(csvSemiColon, csvComma, check.attributes = FALSE)
all.equal(tsv, csvSemiColon, check.attributes = FALSE)
all.equal(fwfWidth, tsv, check.attributes = FALSE)
```





# Save data with readr
##### By re-reading the writes and comparing, we can ensure nothing is lost
```{r}
writeDirectory <- "data/write/"
```

### Values separated by anything
```{r message = FALSE}
anythingPath <- paste(writeDirectory, anythingFile, sep = "")

write_delim(anything, anythingPath, delim = "|")

all.equal(anything, read_delim(anythingPath, delim = "|", trim_ws = TRUE))
```

### Comma separated values
```{r message = FALSE}
commaPath <- paste(writeDirectory, commaFile, sep = "")

write_csv(csvComma, commaPath)

all.equal(csvComma, read_csv(commaPath))
```

### Semi-colon separated values
```{r message = FALSE}
semiColonPath <- paste(writeDirectory, semiColonFile, sep = "")

write_csv2(csvSemiColon, semiColonPath)

all.equal(csvSemiColon, read_csv2(semiColonPath))
```

### Tab separated values
```{r message = FALSE}
tabPath <- paste(writeDirectory, tabFile, sep = "")

write_tsv(tsv, tabPath)

all.equal(tsv, read_tsv(tabPath))
```





# Read Excel data with readxl
```{r message = FALSE}
excelFile <- "words.xlsx"
excelPath <- paste(readDirectory, excelFile, sep = "")

xlsx <- read_excel(excelPath)
xlsx

all.equal(xlsx, anything)
all.equal(xlsx, anything, check.attributes = FALSE)
```





# Read Google Sheets data with googlesheets4
##### googlesheets4 is unable to determine the type of a mixed column (i.e. strings and numerics), and so wraps them in a list. We must explicitly define column data types 
```{r message = FALSE}
# Disable authentication
gs4_deauth()

gooleSheetUrl <- "https://docs.google.com/spreadsheets/d/1xwNqqsZeWfNo5PzDFogusRhRpb0QivuITy558HdqyMM/edit?usp=sharing"

googleSheet <- read_sheet(gooleSheetUrl)
googleSheet

# We can set all column types with a single letter:
# _ or -: skip the column
# ?: guess
# l: logical
# i: integer
# d: double
# D: date
# t: time of day
# T: POSIXct datetime
# c: character
# L: list-column
# C: list of raw cell data

colTypes <- "ccddddddd"

googleSheet <- read_sheet(gooleSheetUrl, col_types = colTypes)
googleSheet

all.equal(googleSheet, xlsx)
```